---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version-ago/starsExtra)](https://cran.r-project.org/package=starsExtra)
[![CRAN_Downloads_Badge](http://cranlogs.r-pkg.org/badges/last-month/starsExtra)](https://cran.r-project.org/package=starsExtra)

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# starsExtra

R package `starsExtra` provides several miscellaneous functions for working with `stars` objects, mainly single-band rasters. Currently includes functions for: 

* Focal filtering
* Detrending of Digital Elevation Models
* Calculating flow length

## Installation

GitHub version:

```{r installation2, eval=FALSE}
install.packages("remotes")
remotes::install_github("michaeldorman/starsExtra")
```

## Usage

Once installed, the library can be loaded as follows.

```{r}
library(starsExtra)
```

## Example

The following code applied a 15\*15 mean focal filter on a 533\*627 `stars` Digital Elevation Model (DEM):

```{r}
data(carmel)
carmel1 = focal2(carmel, matrix(1, 15, 15), "mean")
```

```{r}
data(carmel)
start = Sys.time()
carmel1 = focal2(carmel, matrix(1, 15, 15), "mean")
end = Sys.time()
d = end - start
```

The calculation takes: `r format(d)`.

The original DEM and the filtered DEM can be combined and plotted with the following expressions:

```{r focal-example, fig.width=7, fig.height=4}
r = c(carmel, round(carmel1, 1), along = 3)
r = st_set_dimensions(r, 3, values = c("input", "15*15 mean filter"))
plot(r, breaks = "equal", col = terrain.colors(10), key.pos = 4)
```
