---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version-ago/starsExtra)](https://cran.r-project.org/package=starsExtra)
[![CRAN_Downloads_Badge](http://cranlogs.r-pkg.org/badges/last-month/starsExtra)](https://cran.r-project.org/package=starsExtra)

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# starsExtra

R package `starsExtra` provides several miscellaneous functions for working with `stars` objects, mainly single-band rasters. Currently includes functions for: 

* Focal filtering
* Detrending of Digital Elevation Models
* Calculating flow length
* Calculating the Convergence Index
* Calculating topographic aspect

## Installation

CRAN version:

```{r installation1, eval=FALSE}
install.packages("starsExtra")
```

GitHub version:

```{r installation2, eval=FALSE}
install.packages("remotes")
remotes::install_github("michaeldorman/starsExtra")
```

## Usage

Once installed, the library can be loaded as follows.

```{r}
library(starsExtra)
```

## Documentation

The complete documentation can be found at [https://michaeldorman.github.io/starsExtra/](https://michaeldorman.github.io/starsExtra/).

## Example

The following code applied a 15\*15 mean focal filter on a 533\*627 `stars` Digital Elevation Model (DEM):

```{r}
data(carmel)
carmel_mean15 = focal2(
  x = carmel,             # Input 'stars' raster
  w = matrix(1, 15, 15),  # Weights
  fun = "mean",           # Aggregation function
  na.rm = TRUE,           # 'NA' in neighborhood are removed
  mask = TRUE             # Areas that were 'NA' in 'x' are masked from result
)
```

```{r, include=FALSE}
data(carmel)
start = Sys.time()
carmel_mean15 = focal2(carmel, w = matrix(1, 15, 15), fun = "mean", na.rm = TRUE, mask = TRUE)
end = Sys.time()
d = end - start
```

The calculation takes: `r format(d)`.

The original DEM and the filtered DEM can be combined and plotted with the following expressions:

```{r focal-example, fig.width=7, fig.height=4}
r = c(carmel, carmel_mean15, along = 3)
r = st_set_dimensions(r, 3, values = c("input", "15*15 mean filter"))
plot(r, breaks = "equal", col = terrain.colors(10), key.pos = 4)
```

## Timing

The following code section compares the calculation time of `focal2` in the above example with `raster::focal` (both using C/C++) and the reference method `starsExtra:::focal2r` (using R code only).

```{r}
library(microbenchmark)
library(starsExtra)
library(raster)

data(carmel)
carmelr = as(carmel, "Raster")

res = microbenchmark(
  focal2 = focal2(carmel, w = matrix(1, 15, 15), fun = "mean", na.rm = FALSE), 
  focal = focal(carmelr, w = matrix(1, 15, 15), fun = mean, na.rm = FALSE),
  focal2r = starsExtra:::focal2r(carmel, k = 15, mean),
  times = 10
)
res

```{r focal-timing}
boxplot(res)
```

```{r, include=FALSE}
x = list.files(pattern = "^README-.*\\.png$")
file.copy(x, "docs/", overwrite = TRUE)
```
